---
title: "p8105_hw6_yl5505"
author: "Yan Li"
date: "2024-12-01"
output: github_document
---

```{r, include = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(ggplot2)
library(readxl)
library(patchwork)
library(modelr)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1
```{r}
# Load data using rnoaa::meteo_pull_monitors
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

```
### Bootstraping
```{r}
set.seed(123)

bootstrap_sample = function(data) {
  sample_frac(data, replace = TRUE)
}

bootstrap_results = tibble(bootstrap_id = 1:5000) |>
  mutate(
    bootstrap_sample = map(bootstrap_id, ~ bootstrap_sample(weather_df))
  )

```
### Fit model(with help)
```{r}
bootstrap_results = bootstrap_results |>
  mutate(
    model_results = map(bootstrap_sample, ~ {
      model = lm(tmax ~ tmin, data = .x)
      r_squared = broom::glance(model)$r.squared
      coefficients = broom::tidy(model) |>
        filter(term %in% c("(Intercept)", "tmin")) |>
        pull(estimate)
      
      log_b0b1 = log(prod(coefficients))
      
      tibble(r_squared = r_squared, log_b0b1 = log_b0b1)
    })
  ) |>
  unnest(model_results)

```
### Plot(with help)
```{r}
# Plot R^2 distribution
r2_plot = bootstrap_results |>
  ggplot(aes(x = r_squared)) +
  geom_histogram(binwidth = 0.01, fill = "blue", alpha = 0.6) +
  labs(
    title = "Distribution of R^2",
    x = "R^2",
    y = "Frequency"
  )

# Plot log(b0 * b1) distribution
log_b0b1_plot = bootstrap_results |>
  ggplot(aes(x = log_b0b1)) +
  geom_histogram(binwidth = 0.1, fill = "green", alpha = 0.6) +
  labs(
    title = "Distribution of log(b0 * b1)",
    x = "log(b0 * b1)",
    y = "Frequency"
  )

r2_plot + log_b0b1_plot

```

### 95% CIs
```{r}
ci_r_squared = quantile(bootstrap_results$r_squared, probs = c(0.025, 0.975))
ci_log_b0b1 = quantile(bootstrap_results$log_b0b1, probs = c(0.025, 0.975))

ci_r_squared
ci_log_b0b1

```

# Problem 2

### Data cleaning
```{r}
homicide = read.csv("./homicide-data.csv") |>
  janitor::clean_names() |>
  mutate(
    city_state = paste(city, state, sep = ", "),
    resolved = ifelse(disposition %in% c("Closed by arrest", "Closed without arrest"), 1, 0),
    victim_age = as.numeric(victim_age)
  ) |>
  filter(
    !city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"),
    victim_race %in% c("White", "Black")
  )
```

### logistic regression for Baltimore

```{r}
baltimore = homicide |>
  filter(city_state == "Baltimore, MD")

baltimore_model = glm(
  resolved ~ victim_age + victim_sex + victim_race, 
  data = baltimore, 
  family = binomial()
)

baltimore_results = broom::tidy(baltimore_model) |>
  filter(term == "victim_sexMale") |>
  mutate(
    OR = exp(estimate), 
    CI_lower = exp(estimate - 1.96 * std.error),  
    CI_upper = exp(estimate + 1.96 * std.error) 
  )|>
    select(OR, CI_lower, CI_upper)

baltimore_results

```

### each city
```{r}
city = homicide |>
  group_by(city_state) |>
  nest() |>
  mutate(
    model = map(data, ~ glm(
      resolved ~ victim_age + victim_sex + victim_race, 
      data = .x, 
      family = binomial()
    )),
    results = map(model, ~ broom::tidy(.x) |>
                    filter(term == "victim_sexMale") |>
                    mutate(
                      OR = exp(estimate), 
                      CI_lower = exp(estimate - 1.96 * std.error), 
                      CI_upper = exp(estimate + 1.96 * std.error)
  ) |>
    select(OR, CI_lower, CI_upper))
  ) |>
  select(city_state, results) |>
  unnest(results)

city

```

## PLOT
```{r}

city |>
  arrange(OR) |>
  ggplot(aes(
    y = fct_reorder(city_state, OR),  
    x = OR, 
    xmin = CI_lower, 
    xmax = CI_upper
  )) +
  geom_pointrange() +
  labs(
    title = "Adjusted Odds Ratios (Male vs Female Victims) by City",
    x = "Odds Ratio (Male vs Female Victims)",
    y = "City"
  )
  
```
- The plot shows adjusted odds ratios comparing male and female victims across cities. Cities like Fresno, CA, and Minneapolis, MN, have ORs above 1, indicating higher male victimization odds, while cities like New York, NY, and Long Beach, CA, have ORs below 1, suggesting lower odds for males. Confidence intervals crossing 1 suggest no significant gender difference in some cities.

# Problem 3

### Load and Clean
```{r}
birthweight = read.csv("./birthweight.csv") |>
  janitor::clean_names() |>
  mutate(
    babysex = factor(babysex),
    malform = factor(malform),
    mrace = factor(mrace),
    frace = factor(frace)
  ) |>
  drop_na()

summary(birthweight)
```

## buliding model
```{r}
birthweight_model = lm(
  bwt ~ gaweeks + blength + babysex + wtgain,
  data = birthweight)
 
```
## Plot Residuals
```{r}
birthweight = birthweight |>
  add_predictions(birthweight_model) |>
  add_residuals(birthweight_model)

ggplot(birthweight, aes(x = pred, y = resid)) +
  geom_point(alpha = 0.6) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(
    title = "Residuals vs Fitted Values",
    x = "Fitted Values",
    y = "Residuals"
  ) 
```

* The proposed model predicts birthweight using four key variables: gestational age, birth length, baby’s sex, and mother’s weight gain during pregnancy. These variables were selected based on their biological relevance to birthweight. 
* The model provides a reasonable fit, and shows no major violations of linear regression assumptions.a few outliers suggest that additional refinement may improve its predictive performance.

## Comparsion

### Model 1: Length at birth and gestational age
```{r}
model1 = lm(bwt ~ blength + gaweeks, data = birthweight)

model1_data = summary(model1)|>
  broom::tidy() |> 
  knitr::kable(digits = 3)

model1_data 
```

### Model 2:Head circumference, length, sex, and all interactions (including the three-way interaction)
```{r}
model2 = lm(
  bwt ~ bhead + blength + babysex +
    bhead * blength + blength * babysex + babysex * bhead + 
    bhead * blength * babysex,
  data = birthweight
)

model2_data = summary(model2)|>
  broom::tidy() |> 
  knitr::kable(digits = 3)

model2_data

```

## Cross Validation
```{r}
set.seed(123)
cv_df = crossv_mc(birthweight, 100) |> 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  ) |> 
  mutate(
    my_model = map(train, ~ lm(bwt ~ gaweeks + blength + babysex + wtgain, data = .x)),
    model1_mod = map(train, ~ lm(bwt ~ blength + gaweeks, data = .x)),
    model2_mod = map(train, ~ lm(
      bwt ~ bhead + blength + babysex +
        bhead * blength + blength * babysex + babysex * bhead +
        bhead * blength * babysex, 
      data = .x))
  ) |> 
  mutate(
    rmse_my = map2_dbl(my_model, test, ~ rmse(model = .x, data = .y)),
    rmse_model1 = map2_dbl(model1_mod, test, ~ rmse(model = .x, data = .y)),
    rmse_model2 = map2_dbl(model2_mod, test, ~ rmse(model = .x, data = .y))
  )
```

```{r}
cv_summary = cv_df |> 
  summarise(
    mean_rmse_my = mean(rmse_my),
    mean_rmse_model1 = mean(rmse_model1),
    mean_rmse_model2 = mean(rmse_model2)
  )

cv_summary
```
* Model 2 has the lowest RMSE (288.38), indicating it provides the best predictive accuracy among the three models.
